\chapter{Introduction}
Casual viewers of rowing during the Olympics may not even see a cox
since only 2 of the 14 rowing events that take place are
coxed. 
However, even on a straight course with each boat in its own lane,
coxes are still included in the men's and women's eights races.
Consider next another highly viewed rowing race, the Oxford-Cambridge Boat Race. Coxes are critical since careful steering to stay in the narrow stream of faster water is an important part of this race. The stakes may be higher but even this race is a simpler for a cox than training on a narrow, busy river such as the Cam.

A cox of a college boat on the Cam must deal with coordinating,
coaching and motivating his crew just like the cox of an eight racing
on the rowing lake in the Olympics. He must steer along a narrow and
winding path just like his Boat Race counterparts. On top of this he must also avoid colliding with the many other boats on the river moving in different directions and at different speeds. He will also have a coach on the bank shouting at him to follow a training plan which dictates how fast and how far the boat should travel. And all this must before the crew has to be back for their next lecture with only the rudder to steer the boat and a few commands to the rowers to adjust the speed of the boat. For a new cox this is a lot to deal with. Wouldn't it be great if there were a few, simple rules of thumb he could follow?

The aim of this project is to simulate these conditions on the river Cam representing the crews as autonomous software agents. These agents will then attempt to learn a protocol that a new cox could follow in order to help a crew make the most of its time on the river. To help verify the correctness of the simulation another aim is to create some visualizations for these simulations.

This background report will examine the problem a little more
carefully in the next section. It will then look at some similar work
that might be involved in solving the problem. Finally it will briefly
cover the work that has been done so far and the plan to extend it.
