\appendix
\chapter{Software Engineering Methodology} \label{app:sem}
This project is about more than producing a piece of software but I
consider it to include a significant portion to be software
engineering.  As such I will layout briefly how I plan to follow good
practices. 

\begin{description}
  \item[Version Control] I shall keep my
code (including write-ups like this report) under version control
using Git. Git is very useful to use on individual projects since it
does not require setting up a server for repositories unlike centralized
version control systems such as Subversion. Using Git also allows me
to backup my code on Github where anyone can follow my progress
should they wish, \url{https://github.com/henryaddison}. 
For instance the source for this \LaTeX  report can be found at
\url{https://github.com/henryaddison/msc-background-report}. Of
course, only I will be able to commit changes.

  \item[Iterative] Although the full Agile methodology does not a appropriate to a single-person
project, I still plan to work in an iterative, incremental fashion and
focus on solving one problem at a time, breaking up a complex
task into many, simpler ones. 

  \item[Test Driven Developement] Experience has taught me in the long run, writing
tests is time well spent as it helps to catch and prevent bugs and
also the tests double as a form of living documentation. 

\end{description}

Three months is a very short time so if I do not complete all the aims of the project in
that time I hope these practices will help me write code that I or someone else
could continue to work with at a later date.
